---
title: "BibleSentimentAnalysis"
author: "Nissanka Wickremasinghe"
date: "18 toukokuuta 2016"
output: html_document
---

Sentiment analysis of the bible in English and Finnish

The Bible is the most sold book in the world and has a massive impact on the 
political, social and cutural history of the world. It is also easily available
for free in numerous languages and numerous translations. It does not form a single entity
however, and some versions of the Bible have more books than others. The Ethiopean
bible has the most and the Jewish Bible has the least. It covers a wide range of 
topics, from history and myth all the way down to romantic poetry. This is a first attempt
to study the variation of sentiment and the "shape" of the story of the King James Version 
of the English bible.

Install the required packages
```{r}
library(readr)
library(stringr)
library(syuzhet)
library(dplyr)
library(reshape2)
library(ggplot2)
library(ggthemes)
```


First read the data into R. Skip the first 94 lines of commentary at the top
and used nmax=-1 to read to the end of the file and manually found that 371 lines below are 
commentary so we remove from bottom
```{r}
rawKJV <- read_lines("./pg30.txt", skip = 94, n_max = 114473-371) 
```

Remove all "" marks and remove all those lines
```{r}
noquotesKJV <- rawKJV[nchar(rawKJV)!=0]
```

Create levels for OT and NT - find where they being and end
```{r}
which(grepl("Matthew", noquotesKJV)) #find where the Books start - NT starts at line 63600
```

Create levels for chapter names - find where they being and end
```{r}
which(grepl("Book", noquotesKJV)) #find where each Book start
```

Remove verse numbers and leading spaces and find the locations of where the books
start in the nospaceKJV vector.
```{r}
nonumKJV <- sub("([0-9]+)\\:([0-9]+)\\:([0-9]+)","",noquotesKJV)
nospaceKJV <- c(sub("^\\s+", "", nonumKJV), rep("",6))
bookLocations<- grep("Book", nospaceKJV)
```

In order to get sensible reading of the sentiments, we combined 10 lines of text 
into chunks and perform the sentimet analysis chunk by chunk.
```{r}
chunksKJV <- c(" ")
for (i in seq_along(nospaceKJV)){
  if (i%in% bookLocations){
    chunksKJV[(length(chunksKJV))] <- nospaceKJV[i]
  }
  else{
    if ((i%%10 == 1)&(!(i%in% bookLocations))){
      idx <- length(chunksKJV)+1
      chunksKJV[idx] <- str_c(nospaceKJV[i], 
                                                nospaceKJV[i+1],
                                                nospaceKJV[i+2],
                                                nospaceKJV[i+3],
                                                nospaceKJV[i+4],
                                                nospaceKJV[i+5],
                                                nospaceKJV[i+6],
                                                nospaceKJV[i+7],
                                                nospaceKJV[i+8],
                                                nospaceKJV[i+9], sep = " ")
    }
  }
}
tail(chunksKJV)
head(chunksKJV)
```

We then create a vector which contains the locations of the start of each book in the 
chunksKJV vector
```{r}
BookStart <- c()
for (i in 1:66){
  if (i<10){
    BookStart <- c(BookStart, grep(paste("Book 0", i, sep = ""), chunksKJV))
  }
  else{
    BookStart <- c(BookStart, grep(paste("Book ", i, sep = ""), chunksKJV))
  }
}
BookStart 
```

Create a function that will extract individual books (BookNum must be input by user)
Call extractbook(n = booknumber)
```{r}
extractbook <- function(BookNum){
  if (!BookNum %in% 1:66){
    stop("Enter a number between 1 and 66 inclusive")
  }
  else if (BookNum<=65){
    chunksKJV[BookStart[BookNum]:BookStart[BookNum+1]-1]
  }
  else
    chunksKJV[BookStart[BookNum]:length(chunksKJV)]
} 
```

We use the NRC Word-Emotion Association Lexicon to perform sentiment analysis using the 
"syuzhet" package and create a data frame to store the sentiments (8 sentiment types). 
We then create "testament" factors using the fact that Book 1 to 39 is the Old Testament
and book 40 to 66 is the New Testament.
```{r}
KJVnrc <- cbind(linenumber = seq_along(chunksKJV), get_nrc_sentiment(chunksKJV))
KJVnrc$testament <- "Old Testament"
KJVnrc[BookStart[40]:length(chunksKJV),'testament'] <- "New Testament"
KJVnrc$testament <- as.factor(KJVnrc$testament)
```

We then make a Book name factor for each book and restart a line number for each new book.
```{r}
books <- c()
booknum <- c()
for(i in 1:66){
  books <- c(books,strsplit(chunksKJV[BookStart],"    ")[[i]][[3]])
  booknum <- c(booknum,strsplit(chunksKJV[BookStart],"    ")[[i]][[1]])
}

KJVnrc$book <- " "
for(j in BookStart){
  KJVnrc[j:length(chunksKJV),'book'] <- unlist(strsplit(chunksKJV[j],"    "))[3]
  KJVnrc[j:length(chunksKJV),'booknum'] <- unlist(strsplit(chunksKJV[j],"    "))[1]
}
KJVnrc$testament <- as.factor(KJVnrc$testament)
KJVnrc$book <- as.factor(KJVnrc$book)
KJVnrc$booknum <- as.factor(KJVnrc$booknum)

for(k in books){
  Booklength <- length(KJVnrc$linenumber[KJVnrc$book == k])
  KJVnrc$linenumber[KJVnrc$book == k] <- 1:Booklength
}
```

Overall postive vs. negative sentiment in the text of KJV
```{r}
KJVnrc$negative <- -KJVnrc$negative
```

And then by testament
```{r}
KJVposneg <- KJVnrc %>% select(linenumber, testament, positive, negative) %>%
  melt(id = c("linenumber", "testament"))
names(KJVposneg) <- c("linenumber", "testament", "sentiment", "value")
View(KJVposneg)
```

We first plot the sentiment as points on a chart
```{r}
ggplot(data = KJVposneg, aes(x = linenumber, y = value, color = sentiment)) +
  facet_wrap(~testament, nrow = 3) +
  geom_point(size = 4, alpha = 0.5) + theme_minimal() +
  ylab("Sentiment") + 
  ggtitle(expression(paste("Positive and Negative Sentiment in ", 
                           italic("King James Bible")))) +
  theme(legend.title=element_blank()) + 
  theme(axis.title.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(axis.text.x=element_blank()) +
    scale_color_manual(values = c("aquamarine3", "midnightblue"))
```

The as a barchart
```{r}
ggplot(data = KJVposneg, aes(x = linenumber, y = value, color = sentiment, fill = sentiment)) +
  facet_wrap(~testament, nrow = 3) +
  geom_bar(stat = "identity", position = "dodge") + theme_minimal() +
  ylab("Sentiment") +
  ggtitle(expression(paste("Positive and Negative Sentiment in ", 
                           italic("King James Bible")))) +
  theme(legend.title=element_blank()) +
  theme(axis.title.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(axis.text.x=element_blank()) +
  theme(legend.justification=c(1,1), legend.position=c(1, 0.71)) 
  scale_fill_manual(values = c("aquamarine3", "midnightblue")) +
  scale_color_manual(values = c("aquamarine3", "midnightblue"))
```

The we make a new data frame to calculate the fourier transforms
This uses get_sentiment() which returns numeric vector of sentiment values, 
one value for each input sentence, i.e. the (positive - negative) emotion totals.
```{r}
KJVsentiment <- data.frame(cbind(linenumber = seq_along(chunksKJV), sentiment=get_sentiment(chunksKJV, method = "nrc")))
KJVsentiment$testament <- "Old Testament"
KJVsentiment[BookStart[40]:length(chunksKJV),'testament'] <- "New Testament"
KJVsentiment$testament <- as.factor(KJVsentiment$testament)
```

Plot single sentiment values
```{r}
ggplot(data = KJVsentiment, aes(x = linenumber, y = sentiment)) +
  facet_wrap(~testament, nrow = 2) +
  geom_bar(stat = "identity", position = "dodge", color = "midnightblue") + 
  theme_minimal() +
  ylab("Sentiment") +
  ggtitle(expression(paste("Sentiment in ", italic("King James Bible")))) +
  theme(axis.title.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(axis.text.x=element_blank()) #+
  theme(legend.justification=c(1,1), legend.position=c(1, 0.71))
```

get_transformed_values() - Converts input values into a standardized set of filtered and reverse transformed 
values for easy plotting and/or comparison.
```{r}
KJVft <- as.numeric(get_transformed_values(KJVsentiment$sentiment, 
                                             low_pass_size = 3,
                                             scale_vals = TRUE,
                                             scale_range = FALSE))
KJVft <- data.frame(cbind(linenumber = seq_along(KJVft), ft = KJVft))
View(KJVft)
```

Plot the fourier transforms
```{r}
ggplot(data = KJVft, aes(x = linenumber, y = ft)) +
  geom_bar(stat = "identity", alpha = 0.8, color = "midnightblue", fill = "midnightblue") +
  theme_minimal() +
  ylab("Transformed Sentiment Value") +
  ggtitle(expression(paste("Sentiment in ", italic("King James Bible")))) +
  theme(axis.title.x=element_blank()) +
  theme(axis.ticks.x=element_blank()) +
  theme(axis.text.x=element_blank())  
```